### 2017-12-05 paper/sexSelectionSpacing-ver1.tex
Edit Estimation Strategy section
- Add notes on why not use observed spacing (needs writing)
- Begin rewrite of why not proportionality in hazards
- General edits

% sex-selective Abortion
% Begun originally: 01/11/02
% Begun.: 2017-04-03
% Edited: 2017-04-03

@@

[TK this needs to move to later]
Finally, increased reliability of access and effectiveness of
contraceptive can lead to shorter spacing between births 
\citep{Keyfitz1971,Heckman1976}.
With less reliable contraception parents choose a higher level 
of contraception---resulting in longer spacing---to avoid having 
too many children by accident.
But, as contraception becomes more effective parents can more
easily avoid future births, allowing them to reduce the spacing 
between births without having to worry about overshooting their 
preferred number of children.
This idea may also help explain shorter spacing for better 
educated women than for less educated women, provided that   
knowledge and ability to use contraception differ
across education groups \citep{Tulasidhar1993,Whitworth2002}.

To improve on the literature's current approaches to understanding 
son preference, a new empirical method must be able to capture the 
three main implications of the availability of sex-selective abortions
combined with son preference:
a higher probability that the next child is a son, 
longer spacing to next birth with use of sex selection, 
and that the use of prenatal sex determination may change between 
one observed birth and the next.%
\footnote{
In addition, it should be able to capture the shorter spacing between
births observed in the absence of use or availability of prenatal 
sex determination.
}
This means that we need an empirical model that can jointly estimate 
the association between covariates and both the sex of children born 
and the spacing from last birth. 
My proposed model is a discrete time, non-proportional competing risk 
hazard model with two exit states: either a boy or a girl is born.%
\footnote{
\cite{Merli2000} used a discrete hazard model to examine whether 
there were under-reporting of births in rural China, although they 
estimated separate waiting time regressions for boys and girls.
}

I divide a woman's reproductive life into spells that each covers 
the period between births (from marriage to first birth for the first spell).
For each woman, $i=1,\ldots,n$, the starting point for a spell is time $t=1$ and 
the spell continues until time $t_i$ when either a birth occurs or the spell 
is censored.%
\footnote{
The time of censoring is assumed independent of the hazard rate,
as is standard in the literature.
}
There are two exit states: birth of a boy, $j=1$, or birth of a girl, $j=2$, and 
$J_i$ is a random variable indicating which event took place.
The discrete time hazard rate $h_{ijt}$ is
\begin{equation}
 h_{ijt} = \Pr (T_i=t, J_i=j \mid T_i \geq t; \mathbf{Z}_{it},\mathbf{X}_{i} ),
\end{equation}
where $T_i$ is a discrete random variable that captures when woman $i$'s birth occurs.
To ease presentation the indicator for spell number is suppressed.
The vectors of explanatory variable $\mathbf{Z}_{it}$ and $\mathbf{X}_{i}$ include 
information about various individual, household, and community characteristics 
discussed below.

@@

@@

>>>>

@@

Although it might be tempting to simply examine how observed
spacing has changed over time this is problematic.

Possible ways of doing this:
- Only closed spells (where a birth is observed)
    - can be split by the outcome (ending birth is either male or female)
    - can be split by the sex composition of prior children
- All spells (whether closed or not)
    - can only be split by the sex composition of prior children

Problems:
- Using closed spells only ignore that those families with the strongest
  son preference are also likely to have the longest spacing and therefore be 
  less likely to show up as a complete spell.
- Because of the potential for changes in the use of sex selection within a spell
  it is possible that parents go through multiple sex-selective abortions and still
  end up with a girl
- Some open spells may not result in a birth (childbearing finished with prior birth),
  which is especially a problem for higher parities where sex composition of
  prior children has a large effect on the probability of parity progression.

Need a method that accounts for censoring.

In addition to accounting for censoring, the empirical method must be able 
to capture the three main implications of the availability of sex-selective 
abortions combined with son preference:
a higher probability that the next child is a son, 
longer spacing to next birth with use of sex selection, 
and that the use of prenatal sex determination may change between 
one observed birth and the next.
In addition, it should be able to capture the shorter spacing between
births observed in the absence of use or availability of prenatal 
sex determination.
This means an empirical model that can jointly estimate 
the association between covariates and both the sex of children born 
and the spacing from last birth. 
My proposed model is a discrete time, non-proportional competing risk 
hazard model with two exit states: either a boy or a girl is born.%
\footnote{
\cite{Merli2000} used a discrete hazard model to examine whether 
there were under-reporting of births in rural China, although they 
estimated separate waiting time regressions for boys and girls.
}

I divide a woman's reproductive life into spells that each covers 
the period between births (or from marriage to first birth for the first spell).
For each woman, $i=1,\ldots,n$, the starting point for a spell is time $t=1$ and 
the spell continues until time $t_i$, when either a birth occurs or the spell 
is censored.%
\footnote{
The time of censoring is assumed independent of the hazard rate,
as is standard in the literature.
}
There are two exit states: birth of a boy, $j=1$, or birth of a girl, $j=2$, and 
$J_i$ is a random variable indicating which event took place.
The discrete time hazard rate $h_{ijt}$ is
\begin{equation}
 h_{ijt} = \Pr (T_i=t, J_i=j \mid T_i \geq t; \mathbf{Z}_{it},\mathbf{X}_{i} ),
\end{equation}
where $T_i$ is a discrete random variable that captures when woman $i$'s birth occurs.
To ease presentation the indicator for spell number is suppressed.
The explanatory variable vectors, $\mathbf{Z}_{it}$ and $\mathbf{X}_{i}$, include 
individual, household, and community characteristics discussed below.

@@

The standard proportionality of hazards assumption does not work here.
First, even in the absense of sex selection, it is unlikely that the shape of
hazard functions are identical across groups.
[sex composition]

@@

[TK not sure where to fit this into]
Finally, increased reliability of access and effectiveness of
contraceptive can lead to shorter spacing between births 
\citep{Keyfitz1971,Heckman1976}.
With less reliable contraception parents choose a higher level 
of contraception---resulting in longer spacing---to avoid having 
too many children by accident.
But, as contraception becomes more effective parents can more
easily avoid future births, allowing them to reduce the spacing 
between births without having to worry about overshooting their 
preferred number of children.
This idea may also help explain shorter spacing for better 
educated women than for less educated women, provided that   
knowledge and ability to use contraception differ
across education groups \citep{Tulasidhar1993,Whitworth2002}.
